{\bf Intrusion Detection using Parzen-Windows on Provenance Graph Statistics}

### 2013-12-10 writeup/paper-final.tex
wrote experimental methodology

Below is the list of simple provenance graph statistics we implemented:
\begin{enumerate}
\item number of input and output files
\item number of input and output processes
\item number of input and output pipes
\item number of versions stemming from this node
\end{enumerate}

We also implemented the indegree, Opsahl, and ancestor centrality metrics.
We chose not to implement the Provenance Eigenvector Centrality because computing
the eigenvectors of the large matrix took too much computation time on our machines.
We also chose not to implement age because age would make our data too dependent
on how quickly we executed our workloads.

@@

@@

Because of this exploratory analysis, we chose to implement Opsahl centralty and Ancestory centrality and analyze its accuracy on a real intrusion detection data set.

@@

talk about SPADE and building graphs from BSM

@@

@@

@@

\section{Conclusion}

>>>>

@@

Below is the list of simple provenance graph statistics we implemented:
\begin{enumerate}
\item number of input and output files
\item number of input and output processes
\item number of input and output pipes
\item number of versions stemming from this node
\end{enumerate}
We also implemented the indegree, Opsahl, and ancestor centrality metrics.
We chose not to implement the Provenance Eigenvector Centrality because computing
the eigenvectors of the large matrix took too much computation time on our machines.
We also chose not to implement age because age would make our data too dependent
on how quickly we executed our workloads.

@@

Because of this exploratory analysis, we chose to implement Opsahl and ancestor centrality and analyze its accuracy on a real intrusion detection data set.

@@

We evaluated our approach by analyzing its accuracy on the 1998 DARPA Intrusion Detection Challenge data set (CITE).
The data set consists of users of several weeks of network-based attacks in the midst of normal
background data. The data includes BSM (Basic Security Model) traces (CITE), which provide an
audit of all the system calls made within a certain time period. The data also includes a list
of specific moments in time $T$ and whether an intrusion occurred and if so, what kind of intrusion.

For our experiments, we limited the scope of our intrusion detection system to only u2r attacks, as
the other kinds of attacks were network-based and would not appear in the provenance data. 
We identify each accuracy measurement with a tuple:
$$(\mbox{centrality metric}, \psi, W)$$
where the centrality metric is one of Opsahl, indegree, or ancestor, $\psi$ is our false detection rate for our Parzen-Window
estimates, and $W$ is the time window that we use to form time intervals to look for anomalous nodes. We used
the following procedure to evaluate the accuracy of our system: 
\begin{enumerate}
\item Use BSM traces to build provenance graphs, in a similar fashion to SPADE (CITE). 
\item Compute the centralities of the nodes in the graph using the given centrality metric, and create a mapping from each process
name $P$ to a list of log-likelihood values $L(y)$, where $y$ is the computer centrality measurement for each node in the graph having process name $P$.
\item For a specific moment in time $T$, find all edges in the provenance graph we built with timestamps in the interval $[T - W, T + W]$. 
Use the nodes attached to these edges as our {\em candidate anomaly set}. We used the specific times listed in data sets for our values of $T$.
\item For all nodes in the candidate anomaly set, find the process name $P$ of the node, and use our Parzen-Window approach on process
name $P$ to determine if the node is anomalous. We evaluate whether a node is normal or anomalous using the Parzen-Windows built from the graph
containing the node.
\item For times $T$ in the data set listed as having an intrusion, our intrusion detection system considers that time $T$ as having an intrusion
if there are {\em sufficient} number of nodes in the $[T - W, T + W]$ window considered anomalous. A candidate anomaly set has a 
{\em sufficient} number of anomalous nodes if TODO...
\end{enumerate}

For each $(\mbox{centrality metric}, \psi, W)$, we compute:
\begin{itemize}
\item {\em False Positive Rate (FPR).} Ratio of number of time intervals that do not actually contain an intrusion
but were misidentified as having an intrusion, to the total number of time intervals.
\item {\em True Positive Rate (TPR).} Ratio of the number of time intervals that actually do contain an intrusion
and were correctly identified as having an intrusion, to the total number of time intervals.
\end{itemize}

We evaluated the accuracy of our system on various values of $\psi$ and $W$.

@@

TODO

Need these graphs:
[1]:
- fix psi
- x-axis is W (probably log-scale, as W = 10ms, 1000ms, 100000ms, 10 ** N ms)
- y-axis is [0,1]
- overlay true positive rate for each of the centrality metrics in different color

[2]: repeat [1] but plot false positive rate instead

[3-4] repeat [1-2] with a different value of psi

[5-6] repeat [1-2] with another value of psi

@@

TODO

@@

\section{Conclusion}
TODO
